### 2019-11-12 draft.tex
*** empty log message ***

Let us consider a given set of atoms or molecules, which is composed of
positively and negatively charged particles, and also neutral components.
One can say, correctly, that such a system is an ionized gas, because of
the presence of charged particles in its composition. However, what can
be said about the dominant dynamics of the ionized gas, since neutral and
charged particles have quite a distinct behavior? The neutral particles,
assuming that the gravitational force is negligible, will move freely
in the medium until one collides with another \cite{klimon}. Therefore,
the interaction between the neutral components is characterized as
short-ranged, since it relies only on collisions between two particles.
Charged particles, on the other hand, are sources of electric fields.
Besides that, their motion generates currents that will induce magnetic
fields. The charged particles dynamics is dominated by the presence
of these fields that can act at distance, interacting and  the
fields of several particles at the same time.

@@

On Earth, the plasma is considered an exotic state of matter, associated to the
glowing gas inside fluorescent and neon lamps, microchip manufacturing, research
laboratories, and the promise of cleaner and technically infinite energy source
when scientists and engineers manage to succeed at obtaining a controlled, and
energetically efficient, thermonuclear fusion. Naturally occurring plasmas in our
everyday life are rare, restricted to the occurrence of lightning during storms,
flames and beautiful auroras that are spotted at high latitudes. In the
interplanetary and interstellar space, however, the situation is the opposite.
Starting in the outer layers of Earth's atmosphere, passing through the solar
wind in the interplanetary space, to the stellar interiors and atmospheres,
accretion disks, and molecular clouds it is estimated that something around
$99 \sim 99.9\%$ of the observable matter is in the plasma state \cite{may}.

@@

>>>>

Let us consider a given set of atoms or molecules, which is composed of
positively and negatively charged particles, and also neutral components.
One can say, correctly, that such a system is an ionized gas, because of
the presence of charged particles in its composition. However, what can
be said about the dominant dynamics of the ionized gas, since neutral and
charged particles have quite a distinct behavior? Neutral particles,
assuming that the gravitational force is negligible, will move freely
in the medium until one collides with another \cite{klimon}. Therefore,
the interaction between the neutral components is characterized as
short-ranged, since it relies only on collisions between two particles.
Charged particles, on the other hand, are sources of electric fields.
Besides that, their motion generates currents that will induce magnetic
fields. The charged particles dynamics is dominated by the presence
of these fields that can act at distance, interacting and  the
fields of several particles at the same time.

@@

On Earth, the plasma is considered an exotic state of matter, associated with the
glowing gas inside fluorescent and neon lamps, microchip manufacturing, research
laboratories, and the promise of cleaner and technically infinite energy source
when scientists and engineers manage to succeed at obtaining a controlled, and
energetically efficient, thermonuclear fusion. Naturally occurring plasmas in our
everyday life are rare, restricted to the occurrence of lightning during storms,
flames and beautiful auroras that are spotted at high latitudes. In the
interplanetary and interstellar space, however, the situation is the opposite.
Starting in the outer layers of Earth's atmosphere, passing through the solar
wind in the interplanetary space, to the stellar interiors and atmospheres,
accretion disks, and molecular clouds it is estimated that something around
$99 \sim 99.9\%$ of the observable matter is in the plasma state \cite{may}.

@@

% It is well known that the plasma in our solar system is considered as a natural
% laboratory for the study of space plasma. The wide variety of plasma phenomena
% accessible to direct, in situ measurements and highly resolved remote sensing,
% might be scaled to other astrophysical environments. Besides that, theoretical
% models developed to explain phenomena observed in the solar system plasma might
% be extrapolated to describe plasma process occurring in more extreme astrophysical
% systems.

% Astrophysical plasmas are observed in a vast range of densities, temperatures,
% magnetic fields, and scale lengths\footnote{The description of such an extensive
%   list is out of the scope of this text. On that matter, Ref.~\cite{peratt2014}
%   offers a quite complete exposure of space plasma, and space plasma phenomena
%   occurring in all cosmic scales.}.

% With some exotic exceptions, like the super
% dense and degenerate plasma of white dwarfs (which is beyond the applicability
% of our theory), the solar system hosts a large variety of plasma whose properties

%Plasma is a gas in which the atoms -or molecules- are in an ionized state.
% In such condition, long-range electric and magnetic forces determine the
% interactions of the charged particles. This coupling between the variations
% in the electric and magnetic fields and the motion of ions and electrons
% is the cause of the wide variety of oscillatory processes that characterize
% the plasma dynamics\cite{tsytotp}.

% In this thesis, we present a theoretical model for the inclusion of single
% particle fluctuations and collisional processes related to electrostatic
% oscillation in rarefied, unmagnetized, non-relativistic and fully ionized
% plasmas.

% The problems in solar physics addressed by our theory are related to three
% specific regions of the wide range of action of the solar plasma: solar activity
% regions \cite{Tigik2016a}, transition region between the chromosphere and solar
% corona \cite{Tigik2017a} and the solar wind \cite{Tigik2017b}. The paper

% In this appendix we provide a brief discussion about the space plasma phenomena
% that are addressed by our theory.

% In this text we presented three papers that were published during this Ph.D.
% research, \cite{Tigik2016b,Tigik2017a,Tigik2017b}, and each one of them
% address a different problem in solar physics. There is also another article,
% \cite{Tigik2016a}, which corresponds to the results of the Master's
% dissertation that served as basis for this project.

% Each one of the four papers published during this graduate studies is 
% different solar physics phenomena

% In order to offer a deeper insight on the solar phenomena 
% In this appendix, we make a brief description of the solar processes that are
% being addressed by our theory.

% The Sun is a G2 main-sequence star. From the stellar astrophysics point
% of view, it is considered ordinary, without any unique property that could
% differentiate it from other main-sequence stars; except for the fact that
% it is remarkably close to us. These two facts make the Sun a perfect model
% of study not only for several areas of astrophysics, like stellar structure
% and evolution, but also for plasma physics.

% The theory developed during this graduate research makes use of solar physics
% results to contribute on both sides. In \cite{Tigik2016a},
% \cite{Tigik2016b} and \cite{Tigik2017b}, we present theoretical and
% numerical evidence of plasma processes that might aid in the interpretation
% and accurate modeling  of phenomena observed in the solar plasma. While in
% \cite{Tigik2017a}, we introduce a new equation depicting an unknown process
% and numerically analyze it in the context of solar physics; its existence
% outside theory, however, will depend on the development of more advanced
% space probes to be confirmed.
 ------------------------------------------------------------------------

% Investigations of the origin and behavior of the solar wind, energetic
% particles, and magnetic fields in the heliosphere and their interaction
% with Earth and other planets, as well as with the interstellar medium.
% Investigations of the physics of magnetospheres, including fundamental
% interactions of plasmas and particles with fields and waves, and coupling
% to the solar wind and ionospheres.
% Investigations of the physics of the terrestrial mesosphere, thermosphere,
% ionosphere, including the coupling of these phenomena to the lower atmosphere
% and magnetosphere.

% % The appearance of sunspots is a sign of
% % solar activity, characterized by solar flares and coronal mass ejections.

% This enhanced flux of energetic particles and radiation have direct impact
% in the \emph{space weather} around Earth, triggering geomagnetic storms
% that

% Inside the heliosphere, the Sun, the space environment around Earth and the
% interplanetary medium are extensively explored by space probes.
%% say something about it

% \begin{itemize}
% \item The magnetic dynamo processes involved in the generation and variability
%   of magnetic fields in the Sun, and how it determines the dynamics of the
%   11-year solar cycle of solar activity \cite{Charbonneau2005};
% \item The role of magnetic reconnection in the coronal mass ejection process
%   \cite{Qiu2005};
% \item The coronal heating problem \cite{Vinas2000,PP14,Pezzi2017};
% \item   
% \end{itemize}
 % On that regard, a critical aspect being addressed is the response of Earth’s
 % magnetosphere, ionosphere, and atmosphere to solar inputs.
 % Ref. \cite{schrijver2009}

% \usepackage[inline]{enumitem}
% \begin{enumerate*}[label={\alph*)},font={\bfseries}]
% \item very rarefied plasmas, where collisions can be completely neglected,
%   and
% \item extremely fast phenomena, where the quasi-steady state is attained before
%   collisions can affect the dynamics.
% \end{enumerate*}

% , and have
% been a successful approach for describing kinetic oscillatory instabilities
% and micro-turbulent processes

% As pointed in the introduction, the kinetic theory of plasmas is mostly
% concerned with processes that involve waves and oscillatory instabilities.
% Kinetic instabilities depend on the inter-particle displacement and the
% electrostatic and electromagnetic oscillatory modes that arise as a response.
% This kind of physics cannot be described using a macroscopic approach like
% magnetohydrodynamics (MHD), that treat the plasma as magnetized, incompressible
% fluid. A single-particle description is also out of the question because it
% is both mathematically and computationally prohibitive. Statistical mechanics,
% however, suits very well for the problem. But, as the reader could have noticed,
% the kinetic description of plasmas is quite cumbersome and requires a lot of
% approximations to get to any result. This does not mean that kinetic theory is
% incapable to provide a reliable description of kinetic oscillatory instabilities
% and micro-turbulent processes in real-world plasmas.

 % On Earth, the plasma is considered an exotic state of matter, associated to the
% glowing gas inside fluorescent and neon lamps, microchip manufacturing, research
% laboratories, and the promise of cleaner and practically inexhaustible energy
% source once we have controlled and energetically efficient thermonuclear fusion
% \cite{bell}. Naturally occurring plasmas in our everyday life are rare, restricted
% to the occurrence of lightning during storms, flames, and the beautiful auroras
% that take place at high latitudes. In the space, however, the situation is the
% opposite. Starting in the outer layers of Earth's atmosphere, passing through the
% solar wind in the interplanetary space, to the stellar interiors and atmospheres,
% accretion disks, and molecular clouds it is estimated that something around
% $99 \sim 99.9\%$ of the observable matter in the universe is in the plasma state
% \cite{may}. So understanding the basic physics of plasma processes might help on
% interpreting and describing

% The analysis presented in this doctoral thesis has relevant contributions in the
% study of the combined action of nonlinear kinetic instabilities and collisional
% processes in unmagnetized, fully ionized, hydrogen plasmas. Though the plasma's
% profile under consideration might be considered quite idealized, the physical
% processes that are being addressed have fundamental character - the plasma waves
% under consideration and the particle interactions are inherent to any plasma.
% The underlying nature of these processes means that they might impact the plasma
% dynamics at different astrophysical environments and also in laboratory plasmas.
% Laboratory plasmas, however, are usually constrained by specific experimental
% set-up. This specificity would require in-depth knowledge and maybe access to
% one experiment, which makes the analysis in the experimental context restrictive
% and counterproductive. A more broad approach was then to compare our theoretical
% results with phenomena observed in the naturally occurring space plasma.

% \footnote{In the literature, the term ``astrophysical
%   plasma'' is sometimes used to refer to the plasma observed outside the solar
%   system, while ``space plasma'' is used to account for the plasma in our
%   interplanetary space. In this text, however, both terms are being used as
%   synonyms.}

% \footnote{A comprehensive description of the scales and characteristics of
% cosmic plasmas can be seen in Section 1.2 of Ref. \cite{peratt2014}.}

% The plasma is a collection of ions, electrons and neutral particles that has
% its dynamics dominated by the electric and magnetic fields, generated by the
% electrically charged particles and the electric currents related to their
% motion, respectively. Due to the long-range nature of these self-generated
% fields, one particle affects the movement of several other particles, which
% will interact with other particles, and so on. This \emph{collective behavior}
% \footnote{A detailed exposition of all the processes and conditions that
%   characterize the collective behavior in plasmas, can be accessed in the
%   following introductory books \cite{bitt,gurnett2017,chen,DavidsonNonNeutral}.}
% is what characterizes an ionized gas as a plasma, and is also the reason
% behind the complexity of the plasma dynamics. This complexity comes in the
% form of a myriad of self-consistent oscillatory processes, where the electric
% and magnetic fields affect - and are affected by - the position and motion of
% the charges, and how they vary in time \cite{chen}.

% A complete description of such a complicated dynamics would depend on determining
% the position and momentum of each particle, and their respective microscopic
% electric and magnetic fields, in a given instant in time. This strategy is not
% practical and in most of the cases not feasible for many-body systems, like the
% plasma. However, this microscopic description can be dealt through a hierarchy of
% approximations, leading to the three main theories of plasma physics: kinetic,
% multi-fluid, and magnetohydrodynamics (MHD). The use of each one of these
% formulations will depend on the phenomenon that needs to be described.

Em \cite{klimo} somos apresentados a uma formulação microscópica que combina
os métodos da mecânica clássica para a descrição da variação da posição e do 
momentum das partículas, com as equações de Maxwell para a descrição dos campos 
microscópicos gerados pela movimentação das partículas. Dessa forma, o
estado de um plasma ficaria definido se conseguíssemos determinar as
coordenadas e os momenta de todas as partículas, bem como os valores dos
campos elétrico e magnético em cada ponto, em um determinado instante do
tempo. Com isso, teoricamente, conseguiríamos montar um sistema de equações
que descreveria a evolução temporal do sistema, bastando para isso que
determinássemos as condições iniciais de posição e de momentum para as
partículas e as condições iniciais para os campos.

Contudo, como mencionado anteriormente, essa abordagem é impraticável, devido
a imensa quantidade de partículas que compõem um plasma.
Entretanto, essa formulação pode ser aproximada, permitindo que obtenhamos
uma descrição macroscópica das propriedades do plasma. Existe uma hierarquia
de aproximações que leva às três principais teorias da física de plasmas:
cinética, multi-fluido e magnetoidrodinâmica (MHD). Essas teorias são 
aplicáveis em diferentes contextos, dentro da gama de fenômenos e de condições
que podem ocorrer em um plasma. Como o assunto deste trabalho envolve a
evolução de processos físicos que dependem de detalhes da distribuição de
velocidades das partículas, precisamos aplicar a mais fundamental das três
formulações, a teoria cinética.

A teoria cinética de plasmas consiste na aplicação dos métodos e conceitos da 
mecânica estatística, em conjunto com as equações de Maxwell, para obter uma
descrição macroscópica formal da evolução de um plasma. Este será nosso
ponto de partida.

% This means that a specific charge
% will interact effectively only with other charges that are inside of a
% virtual sphere with radius given by $\lambda_{De}$, being almost completely
% shielded from the influence of particles outside the \emph{Debye sphere}.
% These interactions 
% It is almost because, in the presence of thermal motion, there are small
% fluctuations , creating \emph{Debye shielding}; they are behind the complex
% self-consistent collective behavior of plasmas.
